---
title: "MSDS 410 Assignment 4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(corrplot)
library(purrr)

options(scipen = 999)
```

```{r}
mydata <- data.frame(ames_housing_data)

mydata$TotalFloorSF <- mydata$FirstFlrSF + mydata$SecondFlrSF
mydata$HouseAge <- mydata$YrSold - mydata$YearBuilt
mydata$QualityIndex <- mydata$OverallQual * mydata$OverallCond
mydata$logSalePrice <- log(mydata$SalePrice)
mydata$price_sqft <- mydata$SalePrice/mydata$TotalFloorSF
```

```{r}
subset_df <- mydata %>%
  filter(
    Zoning %in% c("RH", "RL", "RP", "RM", "FV"),
    BldgType == "1Fam",
    SaleCondition == "Normal",
    GrLivArea < 4000,
    Utilities == "AllPub",
    GarageCars >= 1,
    BedroomAbvGr > 0,
    FullBath > 0
  )
```

```{r}
summary(subset_df)
```
```{r}
table(subset_df$TotRmsAbvGrd)
boxplot(subset_df$TotRmsAbvGrd)$out
hist(subset_df$TotRmsAbvGrd)

plot(subset_df$SalePrice~subset_df$HouseAge)
```
```{r}
subset_df <- subset_df %>%
  filter(TotRmsAbvGrd <= 9)
```



```{r}
par(mfrow = c(1, 3))
hist(subset_df$price_sqft)
boxplot(subset_df$price_sqft)
plot(price_sqft~HouseAge, data = subset_df)
```
```{r}
# identify extreme outliers by price_sqft
quartiles <- quantile(subset_df$price_sqft, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(subset_df$price_sqft)
 
ext_lower <- quartiles[1] - 3*IQR
ext_upper <- quartiles[2] + 3*IQR 
 
extr_out_price_sqft <- subset(subset_df, subset_df$price_sqft < ext_lower | subset_df$price_sqft > ext_upper)
extr_out_price_sqft 
```


```{r}
ggplot(data = subset_df, aes(x = Neighborhood, y = SalePrice, group = Neighborhood)) + geom_boxplot(fill = "orange2") +
  ggtitle("Boxplots of SalePrice by NeighborHood") + theme(plot.title = element_text(size = 10, hjust = 0.5)) + 
  theme(axis.text = element_text(angle = 90, vjust=0.5, hjust=1))
```

```{r}
highsale_df <- subset_df %>%
  filter(SalePrice > 400000)
```

```{r}
table(highsale_df$Neighborhood)
```
```{r}
ggplot(data = subset_df, aes(x = Neighborhood, y = GrLivArea, group = Neighborhood)) + geom_boxplot(fill = "orange2") +
  ggtitle("Boxplots of GrLivArea by NeighborHood") + theme(plot.title = element_text(size = 10, hjust = 0.5)) + 
  theme(axis.text = element_text(angle = 90, vjust=0.5, hjust=1))
```

```{r}
table(subset_df$Neighborhood)
```

```{r}
subset(subset_df, subset_df$price_sqft < ext_lower | subset_df$price_sqft > ext_upper)
```
```{r}
subset_df <- subset_df %>%
  filter(SID != 1064)
```

```{r}
nrow(subset(subset_df, subset_df$price_sqft < ext_lower | subset_df$price_sqft > ext_upper))
```
```{r}
hist(subset_df$SalePrice)
```
```{r}
plot(SalePrice~GrLivArea, data = subset_df)
```

```{r}
# identify extreme outliers by sale price
quartiles <- quantile(subset_df$SalePrice, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(subset_df$SalePrice)
 
ext_lower <- quartiles[1] - 3*IQR
ext_upper <- quartiles[2] + 3*IQR 
 
extr_out_SP <- subset(subset_df, subset_df$SalePrice < ext_lower | subset_df$SalePrice > ext_upper)
extr_out_SP 
```
```{r}
hist(subset_df$GrLivArea)
boxplot(subset_df$GrLivArea)
        
hist(subset_df$TotalBsmtSF)
boxplot(subset_df$TotalBsmtSF)

```
```{r}
subset_df %>% filter(GrLivArea > 2800)
```
```{r}
nrow(subset_df %>% filter(GrLivArea > 2800))/nrow(subset_df)
```
```{r}
subset_df <- subset_df %>% filter(GrLivArea <= 2800,
                                  TotalBsmtSF <= 2000)
```



```{r}
subset_df_num <- subset_df %>% select(where(is.numeric))
summary(subset_df_num)
```


```{r}
myfunct2 <- function(x) {
  c(
    "NA Observations" = round(sum(is.na(x)))
    #"Zero Obs" = round(sum(if_else(x == 0, 1, 0)))
  )
}
na_df <- subset_df_num %>%
  map( ~ myfunct2(.)) %>%
  as.data.frame()
row_names <- row.names(na_df)
na_df <- cbind(na_df, row_names) %>%
  gather(key = key, value = value,-row_names) %>%
  spread(key = row_names, value = value)
na_df %>%
  mutate(Variable = key) %>%
  select(Variable, `NA Observations`) %>%
  filter(`NA Observations` > 0)
```
The *MasVnrArea* variable will be imputed with zeros since it's possible that houses do not have that particular feature, imputing a zero will effectively be the same as not having a masonry veneer area. The *LotFrontage* variable will be imputed with the **Median** value. This variable is the linear feet of street connected to the property so it doesn't make sense to impute to zero since properties need to be connected to the street.The NA observations for *BsmtFullBath* and *BsmtHalfBath* will also be imputed with zero since we assume that the missing value means that there is no full or half bath in the basement.

```{r}
subset_df_num <- subset_df_num %>%
  mutate(
    LotFrontage = replace_na(LotFrontage, median(LotFrontage, na.rm = TRUE)),
    MasVnrArea = replace_na(MasVnrArea, 0)
  )
```

```{r}
summary(subset_df_num)
```
```{r}
## Summary table of numeric variables
myfunct3 <- function(x) {
  c("Zero Obs" = round(sum(if_else(x == 0, 1, 0))))
}
zero_df <- subset_df_num %>%
  map(~ myfunct3(.)) %>%
  as.data.frame()
row_names <- row.names(zero_df)
zero_df <- cbind(zero_df, row_names) %>%
  gather(key = key, value = value, -row_names) %>%
  spread(key = row_names, value = value)
zero_df <- zero_df %>%
  mutate(
    "Zero Obs Percent" = scales::percent(`Zero Obs` / nrow(subset_df_num), accuracy = 3L),
    Variable = key
  ) %>%
  select(Variable, `Zero Obs`, `Zero Obs Percent`)
  
zero_df <- zero_df[order(zero_df$`Zero Obs Percent`, decreasing = TRUE), ]
    
knitr::kable(zero_df)
```

```{r}
# drop houses with zero basement sf and with zero kitchens above ground
subset_df_num <- subset_df_num %>%
  filter(TotalBsmtSF > 0,
         KitchenAbvGr >0
         )
```

*HouseAge* is ok to have as 0 since it means it is a new house.

```{r}
subdat <-subset(subset_df_num,
                   select=c("SalePrice", "logSalePrice", "price_sqft",
                            "FirstFlrSF",
                            "TotalBsmtSF", "GrLivArea",
                            "GarageArea", "TotalFloorSF",
                            "LotArea", "LotFrontage",
                            "OverallQual", "QualityIndex",
                            "HouseAge"
                             ))
```

```{r}
data_cor1 <- as.data.frame(
                cor(subdat[ , colnames(subdat) != "SalePrice"],
                subdat$SalePrice))

colnames(data_cor1) <- c('Pearson_Corr_(SalePrice)')

data_cor1 <- cbind(Variable = rownames(data_cor1), data_cor1)
rownames(data_cor1) <- 1:nrow(data_cor1)

data_cor1 <- data_cor1[order(-data_cor1$`Pearson_Corr_(SalePrice)`), ]

knitr::kable(data_cor1)
```
```{r}
data_cor2 <- as.data.frame(
                cor(subdat[ , colnames(subdat) != "logSalePrice"],
                subdat$logSalePrice))

colnames(data_cor2) <- c('Pearson_Corr_(logSalePrice)')
data_cor2 <- cbind(Variable = rownames(data_cor2), data_cor2)
rownames(data_cor2) <- 1:nrow(data_cor2)

data_cor2 <- data_cor2[order(-data_cor2$`Pearson_Corr_(logSalePrice)`), ]

knitr::kable(data_cor2)
```
```{r}
data_cor3 <- as.data.frame(
                cor(subdat[ , colnames(subdat) != "price_sqft"],
                subdat$price_sqft))

colnames(data_cor3) <- c('Pearson_Corr_(price_sqft)')

knitr::kable(data_cor3)
```

```{r}
# remove price_sqft since not meaningful response variable

model_df <-subset(subdat,
                   select=c("SalePrice", "logSalePrice",
                            "FirstFlrSF",
                            "TotalBsmtSF", "GrLivArea",
                            "GarageArea", "TotalFloorSF",
                            "LotArea", "LotFrontage",
                            "OverallQual", "QualityIndex",
                            "HouseAge"
                             ))
```


```{r}
# Summary table of numeric variables
myfunct1 <- function(x) {
  c(
    "SD" = round(sd(x, na.rm = TRUE), digits = 0),
    "Mean" = round(mean(x, na.rm = TRUE), digits = 2),
    "Med" = round(median(x ,na.rm = TRUE), digits = 0),
    "Min" = round(min(x, na.rm = TRUE), digits = 0),
    "Max" = round(max(x, na.rm = TRUE), digits = 0)
  )
}
summary_numeric_df <- model_df %>%
  map( ~ myfunct1(.)) %>%
  as.data.frame()
row_names <- row.names(summary_numeric_df)
summary_numeric_df <- cbind(summary_numeric_df, row_names) %>%
  gather(key = key, value = value,-row_names) %>%
  spread(key = row_names, value = value)
knitr::kable(summary_numeric_df)
```


```{r}
summary(model_df)
```


```{r}
library(scales)

ggplot(model_df, aes(x=TotalFloorSF, y=SalePrice)) +
  geom_point(color="blue", size=2) +
  ggtitle("Scatter Plot of Sale Price vs Total Floor SF") +
  theme(plot.title=element_text(lineheight=0.8, face="bold", hjust=0.5)) +
  geom_smooth(method=lm, se=FALSE, color = "orange") +
  scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma())
```

```{r}
model1 <- lm(SalePrice~TotalFloorSF, data = model_df)

summary(model1)
```

```{r}
anova(model1)
```

```{r}
model_df$y_hat <- 22029.913 + 105.562*model_df$TotalFloorSF

model_df$residual <- model_df$SalePrice - model_df$y_hat

```

```{r}
# standardized residuals and leverage
model_df$leverage1 <- hatvalues(model1)

model_df$std_resid <- rstandard(model1)


# create histogram and scatterplot of standardized residuals
par(mfrow = c(2, 2))
hist(model_df$std_resid)
plot(model_df$y_hat~model_df$std_resid)
```
```{r}
model_df$dffits1 <- dffits(model1)


#find number of predictors in model
p <- length(model1$coefficients)-1

#find number of observations
n <- nrow(model_df)

#calculate DFFITS threshold value
thresh <- 2*sqrt(p/n)

thresh
```


```{r}
model_df[order(-model_df['dffits1']), ]
```
```{r}
#plot DFFITS values for each observation
plot(model_df$dffits1, type = 'h', col = 'blue')

#add horizontal lines at absolute values for threshold
abline(h = thresh, lty = 2)
abline(h = -thresh, lty = 2)
```
```{r}
nrow(model_df %>%
  filter(abs(dffits1) > thresh)
)
```


```{r}
lev_thresh1 <- 3*length(model1$coefficients)/nrow(model_df)

lev_thresh1
```
```{r}
model_df[order(-model_df$leverage1), ]
```
```{r}
#plot leverage values for each observation
plot(model_df$leverage1, type = 'h', col = 'blue')

#add horizontal lines for threshold
abline(h = lev_thresh1, lty = 2)
```

```{r}
lev_inf1 <- model_df %>%
  filter(model_df$leverage1 > lev_thresh1)

nrow(lev_inf1)

```

```{r}
dffits_inf1 <- model_df %>%
  filter(abs(model_df$dffits1) > thresh)

nrow(dffits_inf1)
```


```{r}
model2 <- lm(SalePrice~TotalFloorSF+OverallQual, data = model_df)

summary(model2)
```
```{r}
anova(model2)
```

```{r}
model_df$y_hat2 <- -69506.369 + 60.855*model_df$TotalFloorSF + 26103.013*model_df$OverallQual

model_df$residual2 <- model_df$SalePrice - model_df$y_hat2

model_df$leverage2 <- hatvalues(model2)
model_df$std_resid2 <- rstandard(model2)

# create histogram and scatterplot of standardized residuals
par(mfrow = c(2, 2))
hist(model_df$std_resid2)
plot(model_df$y_hat2~model_df$std_resid2)
```
```{r}
model_df$dffits2 <- dffits(model2)


#find number of predictors in model
p2 <- length(model2$coefficients)-1

#calculate DFFITS threshold value
thresh2 <- 2*sqrt(p2/n)

thresh2
```

```{r}
#plot DFFITS values for each observation
plot(dffits(model2), type = 'h', col = 'blue')

#add horizontal lines at absolute values for threshold
abline(h = thresh2, lty = 2)
abline(h = -thresh2, lty = 2)
```
```{r}
model_df[order(-model_df$leverage2), ]
```
```{r}
lev_thresh2 <- 3*length(model2$coefficients)/nrow(model_df)

lev_thresh2
```
```{r}
#plot leverage values for each observation
plot(model_df$leverage2, type = 'h', col = 'blue')

#add horizontal lines for threshold
abline(h = lev_thresh2, lty = 2)
```

```{r}
lev_inf2 <- model_df %>%
  filter(model_df$leverage2 > lev_thresh2)

nrow(lev_inf2)
```
```{r}
dffits_inf2 <- model_df %>%
  filter(abs(model_df$dffits2) > thresh2)

nrow(dffits_inf2)
```
```{r}
model3 <- lm(SalePrice~TotalFloorSF+OverallQual+GarageArea, data = model_df)

summary(model3)
```
```{r}
anova(model3)
```
```{r}
model_df$y_hat3 <- -71201.392 + 54.811*model_df$TotalFloorSF + 21472.912*model_df$OverallQual + 80.338*model_df$GarageArea

model_df$residual3 <- model_df$SalePrice - model_df$y_hat3

model_df$leverage3 <- hatvalues(model3)
model_df$std_resid3 <- rstandard(model3)

# create histogram and scatterplot of standardized residuals
par(mfrow = c(2, 2))
hist(model_df$std_resid3)
plot(model_df$y_hat3~model_df$std_resid3)
```
```{r}
model_df[order(-model_df$std_resid), ]
```


```{r}
model_df$dffits3 <- dffits(model3)


#find number of predictors in model
p3 <- length(model3$coefficients)-1

#calculate DFFITS threshold value
thresh3 <- 2*sqrt(p3/n)

thresh3
```
```{r}
#plot DFFITS values for each observation
plot(dffits(model3), type = 'h', col = 'blue')

#add horizontal lines at absolute values for threshold
abline(h = thresh3, lty = 2)
abline(h = -thresh3, lty = 2)
```
```{r}
model3_hi <- model_df %>%
  filter(abs(dffits3) > thresh3)
```

```{r}
model3_hi[order(-abs(model3_hi$dffits3)),]
```

```{r}
lev_thresh3 <- 3*length(model3$coefficients)/nrow(model_df)

lev_thresh3
```
```{r}
#plot leverage values for each observation
plot(model_df$leverage3, type = 'h', col = 'blue')

#add horizontal lines for threshold
abline(h = lev_thresh3, lty = 2)
```

```{r}
lev_inf3 <- model_df %>%
  filter(model_df$leverage3 > lev_thresh3)

nrow(lev_inf3)
```

```{r}
dffits_inf3 <- model_df %>%
  filter(abs(model_df$dffits3) > thresh3)

nrow(dffits_inf3)
```
```{r}
model4 <- lm(logSalePrice~TotalFloorSF+OverallQual+GarageArea, data = model_df)

summary(model4)
```
```{r}
anova(model4)
```
```{r}
model_df$y_hat4 <- 10.67989850 + 0.00031010*model_df$TotalFloorSF + 0.11754934*model_df$OverallQual + 0.00039333*model_df$GarageArea

model_df$residual4 <- model_df$logSalePrice - model_df$y_hat4

model_df$leverag4 <- hatvalues(model4)
model_df$std_resid4 <- rstandard(model4)

# create histogram and scatterplot of standardized residuals
par(mfrow = c(2, 2))
hist(model_df$std_resid4)
plot(model_df$y_hat~model_df$std_resid4)
```



```{r}
model_df$dffits4 <- dffits(model4)


#find number of predictors in model
p4 <- length(model4$coefficients)-1

#calculate DFFITS threshold value
thresh4 <- 2*sqrt(p4/n)

thresh4

```

```{r}
#plot DFFITS values for each observation
plot(dffits(model4), type = 'h', col = 'blue')

#add horizontal lines at absolute values for threshold
abline(h = thresh4, lty = 2)
abline(h = -thresh4, lty = 2)

```

```{r}
model4_hi <- model_df %>%
  filter(abs(dffits4) > thresh4)
```

```{r}
model4_hi[order(-abs(model4_hi$dffits4)),]
```


```{r}

model3_rmse <- sqrt(mean(model3$residuals ^ 2))
model4_rmse <- sqrt(mean(model4$residuals ^ 2))
model3_summary <- summary(model3)
model4_summary <- summary(model4)

# Summary Table
summary_df <- tibble(
  "Model" = c(3, 4),
  "R-Squared" = c(
    model3_summary$r.squared,
    model4_summary$r.squared
  ),
  "Adjusted R-Squared" = c(
    model3_summary$adj.r.squared,
    model4_summary$adj.r.squared
  )
)
# Output
knitr::kable(summary_df)
```
```{r}
boxplot(model_df$SalePrice)
```
```{r}
# create df with obs under dffits limit
model_df_no_outlier_dffits4 <- model_df %>%
                        filter(abs(model_df$dffits4) < thresh4)


boxplot(model_df_no_outlier_extr$SalePrice)
```

```{r}
model4_no_out_d <- lm(logSalePrice~TotalFloorSF+OverallQual+GarageArea, data = model_df_no_outlier_dffits4)

summary(model4_no_out_d)
```
```{r}
# number of high influential points removed DFFITS
nrow(model_df)-nrow(model_df_no_outlier_dffits4)

```
```{r}
model4_no_out_d_rmse <- sqrt(mean(model4_no_out_d$residuals ^ 2))
model4_dffits <- summary(model4_no_out_d)

# Summary Table
summary_df2 <- tibble(
  "Model" = c(4, '4_dffits'),
  "Observations Removed" = c(0, nrow(model_df)-nrow(model_df_no_outlier_dffits4)  ),
  "R-Squared" = c(
    model4_summary$r.squared,
    model4_dffits$r.squared
  ),
  "Adjusted R-Squared" = c(
    model4_summary$adj.r.squared,
    model4_dffits$adj.r.squared
    
  )
)
# Output
knitr::kable(summary_df2)
```
```{r}
model5 <- lm(SalePrice~TotalFloorSF+OverallQual+GarageArea+TotalBsmtSF+LotArea+HouseAge, data = model_df)

summary(model5)
```

```{r}
anova(model5)
```

```{r}
model_df$y_hat5 <- -48836.2117 + 52.9681*model_df$TotalFloorSF + 15803.3495 *model_df$OverallQual + 
                    41.9937*model_df$GarageArea + 35.0784*model_df$TotalBsmtSF + 0.8671*model_df$LotArea 
                    - 328.7413*model_df$HouseAge
                    
model_df$residual5 <- model_df$SalePrice - model_df$y_hat5

model_df$leverage5 <- hatvalues(model5)
model_df$std_resid5 <- rstandard(model5)

# create histogram and scatterplot of standardized residuals
par(mfrow = c(1, 2))
hist(model_df$std_resid5)
plot(model_df$y_hat5~model_df$std_resid5)
```
```{r}
model_df[order(-model_df$std_resid5),]
```

```{r}
plot(model_df$SalePrice~model_df$std_resid5)
plot(model_df$GrLivArea~model_df$std_resid5)
```


